---
title: "STAT 210 Lab 7: Murder in SQL City"
author: "Ericka Quintero"
date: "18 April 2025"
format: html
embed-resources: true
execute: 
  warning: false
  message: false
---
For this lab, you will be joining and filtering related data sets to solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that says 
`"Your name here!"`. Change this to be your name and save your file either by
clicking on the blue floppy disk or with a shortcut (command / control + s). 

Be sure to [commit the files to your repo](https://happygitwithr.com/existing-github-first#stage-and-commit). 

Let's get started!


# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid "object junk"
in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users' database
skills, so they created a murder mystery. Can you solve this crime in SQL City??

The relational data you will be working with contains tables with different
pieces of information pertinent to the crime - people, social media check-ins,
driver's licenses, crime scene reports, police interviews, and more!


## Access the Data

This code chunk will read in **all** of the tables of data for you. Don't modify or
remove this! Take some time to look at each file type so that 

```{r}
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in SQL City sometime on January 15, 2018. Could you retrieve the crime scene report from the police department’s database and follow the clues to find the person responsible for the murder?!


**Step 1: Find the police report in `crime_scene_report`. Then used the information in the report to move on to the next data source based on the information learned.**

```{r}
#| label: crime scene report in SQL City
# Code for looking at the relevant crime scene report.
sqlmurder_report <- crime_scene_report %>%
  filter(date == "20180115",
         type == "murder",
         city == "SQL City",
         )
(sqlmurder_report)

```
Description:
Security footage shows that there were 2 witnesses. The first witness lives at the last house on "Northwestern Dr". The second witness, named Annabel, lives somewhere on "Franklin Ave".

Notes:
  First witness is identified as Morty Schapiro (14887), who lives on 4919 Northwestern Dr (the last house on the street). Mr. Schapiro describes on the interview that he heard the gun shot and saw the suspect ran out. Suspect is described wearing a "Get Fit Now" gym bag with a membership number starting with "48Z". Witness states that only gold members are allowed to have those. License plate begins with "H42W".
    Second witness is identified as Annabel Miller (16371), who lives on 103 Franklin Ave. Ms. Miller describes on the interview that she saw the murder occur and recognizes the murderer from a gym session dating to last week on January 9, 2018.
    After reviewing Get Fit Now records, there are two possible suspects that share "48Z" in there gold membership status: Joe Germuska (28819) and Jeremy Bowers (67318). Jeremy attended "Funky Grooves Tour" the day of the murder.



**Next Steps: Follow the evidence to the person responsible for the murder, building a report as you go.**
There are accomplices, some knowingly and some unknowingly, but there is only one mastermind.
```{r}
#| label: file on witness one
witness_one <- person %>%
filter(address_street_name == "Northwestern Dr") %>%
arrange(desc(address_number)) # arranges address numbres in descending order

interview %>%
  filter(person_id == "14887")

(witness_one)

```

```{r}
#| label: file on witness two
witness_two <- person %>%
filter(name == "Annabel Miller",
       address_street_name == "Franklin Ave")

interview %>%
  filter(person_id == "16371")

(witness_two)
```

```{r}
#| label: suspect case file

suspect <- get_fit_now_check_in %>%
  filter(check_in_date == "20180109" &
  substr(membership_id, 1,3)  == "48Z") # extracts variables that include "48z in id

get_fit_now_member %>%
  filter(membership_status == "gold" &
  substr(id, 1, 3) == "48Z"
         )
drivers_license %>%
  filter(gender == "male" &
  str_detect(plate_number, "H42W")) # detects the presence of "H42W" in a licence plate

facebook_event_checkin %>%
  filter(date == "20180115",
         person_id == "28819" |
         person_id == "67318")
```

```{r}
#| label: suspect licence identification
 
licence_ids <- person %>%
  filter(name == "Jeremy Bowers" |
         name == "Joe Germuska")

drivers_license %>%
  filter( id == "173289" |
          id == "423327")

licence_ids
```




Solve the murder mystery, showing **all of your work in this document**. Your
document and code must be well organized, easy to follow, and reproducible.

+ Use headers and written descriptions to indicate what you are doing.
+ Use `dplyr` verbs and `join` functions rather than just looking through the
tables manually. Functions from `stringr` and `lubridate` will also be useful.
+ Use good code formatting practices.
+ Comment your code.
+ Cite any external sources you use to solve the mystery.



::: callout-note
## And the final suspect is...

**Jeremy Bowers**
:::
